---
title: "Assignment 1: Analyzing an Obesity dataset"
author: William Schaafsma, Hau Nguyen, Qingyu Meng
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
       highlight: textmate
       theme: flatly
       number_sections: TRUE
       toc: TRUE
       toc_float:
         collapsed: TRUE
         smooth_scroll: FALSE
---

# Introduction and Research Question

In this assignment an obesity dataset, retrieved from [Palechor & Manotas, 2019](https://www-sciencedirect-com.proxy.library.uu.nl/science/article/pii/S2352340919306985?via%3Dihub) will be analyzed. The data gathered for this paper comes from individuals that live in Peru, Mexico and Colombia. It's stated that 23% of the data is "real" and that 77% has been synthetically generated.

The goal for this assignment is to build a model that most accurately predicts the BMI of *new* patients. Understanding which predictors have great influence for one's BMI is favorable for policymakers regarding national health. Moreover, patients get more understanding of their own health.

We will work in an exploratory manner. Thus, in this assignment we are particularly interested in: 'What predictors show the greatest influence regarding a person's BMI?'

# Packages

The following libraries are required to run this program succesfully.
```{r libraries, warning = FALSE, message = FALSE}
library(ggplot2)
library(corrplot)
library(dplyr)
library(knitr)
library(tidyverse)
library(mice)
library(magrittr)
library(readr)
library(caret)
```


# Loading the dataset & set seed

First we load the data and set the seed in order to generate a reproducible result.

```{r, message=FALSE}
# loading data
obese <- read_csv("ObesityDataSet_raw_and_data_sinthetic (2)/ObesityDataSet_raw_and_data_sinthetic.csv")

# set the seed
set.seed(1705)
```

# Processing of the data

We want to include BMI as a dependent variable in our model. Thus we have to derive BMI from Weight and Height with the formula:  $BMI = Weight / Height^2$

```{r, message=FALSE, }
# calculating bmi from height and weight
Bmi <- data.frame(Bmi = (obese$Weight / (obese$Height^2)))

# adding bmi to the original dataframe
obese_complete <- cbind(obese, Bmi)
```


# Splitting the data into train, validation & test

In order to evaluate the performance of the model we will build later on, we have to split the data into three partitions: Train, Test, Validation. The data will be split in a way that ~70% of the data is used for training the model, ~20% for validating the model and ~10% for testing the model.

```{r, message=FALSE}
# partition of training
part_train <- createDataPartition(obese_complete$Bmi, p = .7, 
                                  list = FALSE, 
                                  times = 1)

# creation of training data
obese_train <- obese_complete[part_train,]

# remainder for validation and testing
part_test_val <- obese_complete[-part_train,]

# partition of validation
part_val <- createDataPartition(part_test_val$Bmi, p = .66, 
                                  list = FALSE, 
                                  times = 1)
# creation of validation data
obese_val <- part_test_val[part_val,]

# creation of test data
obese_test  <- part_test_val[-part_val,]
```

# Explaining the variables

In this section, an elaboration on the variables is given. For starters, the categorical variable *Gender* consists of two categories representing one's gender: male or female.

Next is the continuous variable *Age* that only contains integers of every person's age. 

Thenthe continuous variable *Height* that contains floating values for a person's Height, measured in meters.

Then there's another continuous variable *Weight* that also contains floating values for a person's weight, measured in Kilograms (KG).

**Further work on this!!!** 



# Getting to know the data

In order to work with a dataset, we need to understand the dataset. First, let's check wether there are missing data in our dataset. To do this we use the `mice` package.

```{r, message=FALSE}
md.pattern(obese_complete)
```

Fortunately, we can conclude that our dataset contains no missing values. Furthermore, the figure shows that our dataset holds 18 variables and 2111 observations. 

Now, let us see the head and the summary of the dataset

```{r head}
head(obese_complete) %>% 
  knitr::kable(format = "markdown", digits= 1, padding = 30, align = 'c')
```

```{r}
summary(obese_complete)
```


# Correlations

# Model training

# Best model selection & model comparison

# Model evaluation & VIF

# Checking assumptions

# Visualizations

# Answering research question

# Discussion & Limitations

# Sources
